A new method to estimate dark matter halo concentrations

### 2015-10-19 paper/article.tex
primera version completa

\begin{figure*}
\begin{center}
  \includegraphics[width=0.90\textwidth]{concentration.pdf}
\end{center}
\caption{Mass-concentration relationship for the three different
  methods used on the same cosmological N-body data. The central lines
  correspond the median and the shadowed region indicates the
  quartiles.
    \label{fig:concentration}}
\end{figure*}

\section{Conclusions}
\label{sec:conclusions}

In this paper, we presented a new method to estimate the
concentration of dark matter halos in N-body simulations. 
We tested our method on mock halo data to study the impact of total
number of particles and input concentration on the retrieved values.
We compared these results against two other methods commonly used in
the literature to estimate concentrations.
Finally, we applied our method to halos extracted from a cosmological
N-body simulation to estimate the impact of our method on the mass
concentration relationship.

Comparison against other methods.\\

The effect of the number of particles.\\

The effect of the input concentration.\\

The impact on the mass-concentration relationship.\\

>>>>

\begin{figure*}
\begin{center}
  \includegraphics[width=0.75\textwidth]{concentration.pdf}
\end{center}
\caption{Mass-concentration relationship for the three different
  methods used on the same cosmological N-body data. The central lines
  correspond the median and the shadowed region indicates the
  quartiles.
    \label{fig:concentration}}
\end{figure*}

### 2015-10-19 paper/paper.tex
primera version completa

>>>>

\label{sec:data}
\begin{figure*}
  \begin{center}
    \includegraphics[width=0.33\textwidth]{mass-velocity.pdf}
    \includegraphics[width=0.33\textwidth]{mass-density.pdf}
    \includegraphics[width=0.33\textwidth]{density-velocity.pdf}
  \end{center}
  \caption{Comparison between the concentrations measured by our
    method and the maximum velocity (left) and density (middle)
    methods. The line indicates the equal value between the two
    techniques. The right panel compares the results of the maximum
    velocity and density methods.
  \label{fig:mdv}}
\end{figure*}

We use data from the MultiDark cosmological simulation that follows
the non-linear evolution of a dark matter density field sampled with
$2048^3$ particles over a cubic box of $1000$ \hMpc on a side. 
The data is publicly available through \url{http://www.cosmosim.org/}.
More details about the structure of the database and the simulation
can be found in \citep{2013AN....334..691R}.

We build a sample of all halos located in a cubic sub-volume of $100$
\hMpc on a side centered on the most massive halo in the simulation at
$z=0$ which corresponds to the \texttt{miniMDR1} tables in the
database.
From this sample we select all the halos at $z=0$ detected with a
Friends-of-Friends (FoF) algorithm with masses in the interval
$10^{12}\leq M_{\rm FoF}/\hMsun \leq 10^{15}$.
The FoF algorithm used with a linking length of $0.17$ times the average
inter-particle distance. This choice translates into an overdensity
$\Delta_h\sim 400-700$ dependent on the halo concentration
\citep{More2011}.
Finally, we select from the database all the particles that belong to
each FoF group.

From this set of particles we follow the procedure spelled out in
Section \ref{sec:method} with $\Delta_h=740$  (corresponding to $200$
times the critical density) to select an spherical region that we
redefine to be our halo.
This choice makes that our overdensities are fully included inside the
original FoF particle group.
On the interest of providing a fair comparison against the density
method we only report results from overdensities with at least $200$
particles.

Figure \ref{fig:concentration} shows the results of the
mass-concentration relationship.

\begin{figure*}
\begin{center}
  \includegraphics[width=0.75\textwidth]{concentration.pdf}
\end{center}
\caption{Mass-concentration relationship for the three different
  methods used on the same cosmological N-body data. 
  The lines correspond to the median concentration values in each bin.
    \label{fig:concentration}}
\end{figure*}

\section{Conclusions}
\label{sec:conclusions}

In this paper, we presented a new method to estimate the
concentration of dark matter halos in N-body simulations. 
We tested our method on mock halo data to study the impact of total
number of particles and input concentration on the retrieved values.
We compared these results against two other methods commonly used in
the literature to estimate concentrations.
Finally, we applied our method to halos extracted from a cosmological
N-body simulation to estimate the impact of our method on the mass
concentration relationship.

The first benchmark was performed on mock halos generated with known
concentration values for different particle numbers. 
For all methods, the accuracy in retrieveing the input concentration
increases with the number of particles as summarized in Figure \ref{fig:error}.
Our method systematically shows smaller errors in the retrieved values
compared to the other two methods, going from $10\%$ for halos with
$20$ particles down to $0.01\%$ for halos with $2\times 10^4$ particles. 
In our method the average error on the retrieved values decreases with
the number of particles as $\langle|D|\rangle\propto N^{-1}$.
The error in density method initially behaves as $\langle|D|\rangle\propto
N^{-3/2}$ but at $N=2\times 10^3$ it saturates to $0.1\%$. 
For the velocity method the error behaves closely to
$\langle|D|\rangle\propto N^{-1/2}$.
We also gauged the effect of different concentration parameters at
fixed particle numbers. 
We found larger offsets for larger input concentrations. 
The offset goes in the sense that the measured concentration tends to
be underestimated.
Nevertheless, the dominant effect on the measurement is the particle
number.

We used a N-body cosmological simulation to test the impact of our
method on the mass-concentration relationship.
Although the three methods are in general agreement there are
noticeable differences in the mass concentration relationship.
